---
title: "Clustering"
author: "Sean F. Larsen"
date: "December 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("cluster")
library("rattle.data")
library("NbClust")
```

## Clustering
This project is based on the K-Means exercise from 'R in Action'
For the original blog post and solutions go to  
<http://www.r-bloggers.com/k-means-clustering-from-r-in-action/>

I will be using the cluster, rattle.data, and NbClust packages.
I will also be usinging the Wine Data set from the rattle.data package.
This code will pull the data and show me the column heads for further analysis.

```{r echo=FALSE}
data(wine, package="rattle.data")
head(wine)
```
For this exercise I will remove the first column from the data
and the scale the data.

```{r echo=FALSE}
wine$Type <- NULL
head(wine)
wine.scale <- scale(wine, center = TRUE, scale = TRUE)
```

Now we are going to cluster and plot the data to determine the number of clusters.

```{r echo=FALSE}
set.seed(1234)
nc <- NbClust(wine.scale, min.nc=2, max.nc=15, method="kmeans")
barplot(table(nc$Best.n[1,]),
	          xlab="Numer of Clusters", ylab="Number of Criteria",
		            main="Number of Clusters Chosen by 26 Criteria")
```

Based on the results I am going to select 14 clusters.

## Run the K-Means and Plot
Now that I have picked the number of clusters, I'm running k-means 
using this number of clusters. This code will output the result of calling kmeans()
into a variable fit.km
Then we will plot the data.

```{r echo=FALSE}
fit.km <- kmeans(wine.scale, 14, iter.max = 10, nstart = 1)
table(fit.km$cluster)
clusplot(wine.scale, fit.km$cluster)
```


